% Christina
First and foremost, I would like to express my sincere gratitude to \emph{Dr. Christina Ludwig} for the opportunity of conducting this research, her continuous support, guidance, and insightful discussions that have tremendously enriched this work. Your dedication and knowledge have been invaluable to this study and my academic journey in general.

% Sven
I extend my heartfelt thanks to \emph{apl. Prof. Dr. Sven Lautenbach} for his valuable insights and constructive feedback throughout this research. Your expertise has greatly contributed to the development of this study, including the refinement of the research questions and the overall aim.

% Maciej
Special thanks go to my colleague, \emph{Dr. Maciej Adamiak}, for his assistance, provision of the utilized framework, and inspiration that lay the foundation of this work. Without your limitless inventiveness, creativity, and passion, this thesis would have not been conducted so successfully. Thanks also for enhancing my understanding of the subject matter during the Innovation Weeks and our multiple meetings, this helped me greatly for this thesis.

% Institute of Geography & Heigit
I am grateful to the \emph{Institute of Geography} and the \emph{HeiGIT gGmbH} for providing the resources and environment necessary for the completion of this thesis. 

% Friends and Family
I thank my \emph{family} for continuous support and encouragement throughout my studies. You always had my back and spanned a boundless safety net, enabling me to focus on my studies. Thanks also go to my \emph{friends} for multiple joyful and insightful discussions and experiences.

% Sara
Thanks also to you, \emph{Sara}, without you I would not have left my desk to enjoy the sun in the summer, take relaxing walks, and the office and us would miss a very welcome distraction.

% Bu
Finally, I would like to express my deepest gratitude to \emph{my better half} for unwavering and enduring support, patience, and the best food I could ask for during the long final sprint. Your love and belief in me have been a constant source of motivation and a large driver for my recent development in any aspect.

% Final
Thank you all for giving me the opportunity to grow, learn, and follow my dreams.  Without you, this would not have been possible.